Opt-K

### 2013-03-15 paper/paper.tex
stuff
change
working better...
changed notation
notation shift
power
Added figures to the paper. I think I have too many in there so we'll have to decide what to keep.

\begin{equation}
\label{eq:like}
P[x_{0...T-1}] dx^N = \prod_{i=0}^{T-1} T(x_i \rightarrow x_{i+1}) \cdot \prod_{i=0}^{T} p(x_{i} | \sigma(x_{i}))
\end{equation}

@@

Briefly, we generated uniform random vectors within a bounding box whose size was slightly larger than the data. This random sample was then accepted if it was 'close enough' to a generator in the 100 state model, where close enough meant it was within 0.2 units from any generator in that model. This was used to approximate the convex hull of the data. Then the accepted points were assigned to a state. Then the volume of a state is proportional to the number of samples assigned to that state divided by the total number of samples. We note, that these volumes are actually in units of the total acceptable volume. This affects the absolute value of the likelihood, however, since the total volume was constant in all cases, the relative likelihoods are the same.
\begin{figure}[h!]
\includegraphics[width=3in]{figs/muller_pot_traj.pdf}
\caption{The M\"{u}ller potential consists of three metastable wells. We simulated Langevin dynamics on this potential as a validation of the new likelihood function.}
\label{fig:muller_pot}
\end{figure}

@@

\begin{equation}
BIC = -2 \log L + m \log(n) 
\label{eq:bic}
\end{equation} Here, $L$ denotes the likelihood of a model, while $m$ is the number of parameters used in the model, and $n$ is the number of data points. Essentially this is a mathematical way of saying we would like to use a model with the maximum likelihood, but without introducing too many parameters. 
% don't know how to say this yet:
%We note that we cannot take this too seriously, and there are many other such criteria, such as the Aikaike Information Content (AIC). 
We found that the BIC of the models built using K-Centers clustering has a minimum at $k=250$. This minimum is the ``optimal'' model according to the BIC of our likelihood. As a check of the usefulness of this minimum, we plotted the three slowest implied timescales in the models as a function of $k$. The optimal model was close to the point at which the implied timescales became constant.

@@

\bibliography{bibliography}
\end{document}

>>>>

\begin{equation}
P[X_{0...T-1}] dx^T = \prod_{i=0}^{T-1} T(X_i \rightarrow X_{i+1}) \cdot \prod_{i=0}^{T} p(X_{i} | \sigma(X_{i}))
\label{eq:like}
\end{equation}

@@

Briefly, we generated uniform random vectors within a bounding box whose size was slightly larger than the data. This random sample was then accepted if it was 'close enough' to a generator in the 100 state model, where close enough meant it was within 0.2 units from any generator in that model. This was used to approximate the convex hull of the data. Then the accepted points were assigned to a state. Then the volume of a state is proportional to the number of samples assigned to that state divided by the total number of samples. We note, that these volumes are actually in units of the total acceptable volume. This affects the absolute value of the likelihood, however, since the total volume was constant in all cases, the relative likelihoods are the same.
\begin{figure}[h!]
\includegraphics[width=3in]{figs/muller_pot.png}
\caption{The M\"{u}ller potential consists of three metastable wells. We simulated Langevin dynamics on this potential as a validation of the new likelihood function.}
\label{fig:muller_pot}
\end{figure}

@@

\begin{equation}
BIC = -2 \log L + m \log(n) 
\label{eq:bic}
\end{equation} Here, $L$ denotes the likelihood of a model, while $m$ is the number of parameters used in the model, and $n$ is the number of data points. Essentially this is a mathematical way of saying we would like to use a model with the maximum likelihood, but without introducing too many parameters. 
% don't know how to say this yet:
%We note that we cannot take this too seriously, and there are many other such criteria, such as the Aikaike Information Content (AIC). %%%
We found that the BIC of the models built using K-Centers clustering has a minimum at $k=250$. This minimum is the ``optimal'' model according to the BIC of our likelihood. As a check of the usefulness of this minimum, we plotted the three slowest implied timescales in the models as a function of $k$. The optimal model was close to the point at which the implied timescales became constant.

@@

\begin{figure*}
\includegraphics[width=5in]{figs/kcent_vors.png}
\caption{For K-Centers models, we observed a rapid increase in likelihood followed by a plateau. For models built with too few states, the states within the metastable wells are too large. By increasing the number of states, the resolution in increased, but using too many states leads to a model like the bottom right which is clearly over-fit and so has a non-optimal BIC despite having a higher likelihood.}
\end{figure*}

\begin{figure}
\includegraphics[width=3in]{figs/cluster_comp.png}
\caption{We also compared three clustering algorithms and used the likelihood function to determine which is best. The likelihood of the K-Means models was always larger than the other two. Additionally, the K-Means algorithm plateaued at a lower value of $k$, which corresponded to a minimum in the BIC curve at 200 states vs. the optimal models of 250 states in the K-Centers and Hybrid K-Medoids models.} \end{figure}

\begin{figure}
\includegraphics[width=2in]{figs/clust_vors.png}
\caption{The likelihoods (and BIC's) of the three clustering algorithms predicted that the K-Means algorithm is the better clustering algorithm. The result of K-Means is many states clustered into the highest density areas of the potential. The Hybrid K-Medoids ends up somewhere in between the K-Centers model and the K-Means model. This is likely due to the implementation being an approximate algorithm that requires more iterations to converge to the best state decomposition.}
\end{figure}
\bibliography{bibliography}
\end{document}
